---
title: "SI Code"
date:  '`r format(Sys.time(), "%d %B, %Y")`'
author: "Chris Terry"
output:
  github_document:
    html_preview: false
editor_options: 
  chunk_output_type: console
---

## Introduction

This is a tidied version of the code used to generate the supplementary results in Terry et al (2019) *Interaction modifications lead to greater robustness than pairwise non-trophic effects in food webs*

```{r, message=FALSE, warning=FALSE}
library(deSolve)
library(igraph)
library(numDeriv)
library(rootSolve)
library(NetIndices)
library(pbapply)
library(knitr)
library(cowplot)
library(mvbutils)
library(Rcpp)
library(pracma)
library(parallel)
library(tidyverse)

NUM_CORES <- 3 # Number of cores to use for parallel

```

```{r, message=FALSE, warning=FALSE}
try(sapply(paste0('Scripts/',list.files('Scripts')), source)->tmp)

ParallelRun <- function(DATA,FUNC,Name, ...){
  c1<- makeCluster(NUM_CORES)
  clusterEvalQ(c1,
               {require(dplyr);require(rootSolve);
                 require(deSolve);require(igraph);require(purrr)})->tmp
  clusterExport(c1,as.list(find.funs()[find.funs()!='LV']))
  pblapply(X =DATA, cl = c1,FUN = FUNC,...)-> tmp
  assign(value = tmp,Name)
  save(list=Name,file=paste0('Output from Simulations/',Name))
  stopCluster(c1) 
}

Unpacker <- function(SafeObject){
  SafeObject  %>%transpose %>% `[[`('result') %>% compact() -> x
  return(x)
}

ErrorFinder<- function(SafeObject){
  SafeObject  %>%transpose %>% `[[`('error') -> errormessages
  errormessages %>% map_lgl(is.null) %>% `!` %>% which %>% print
  errormessages %>% compact() %>% print()
}
```

# SI 

## SI 1. Body mass distribution

```{r}
load('Output from Simulations/BaseCommunities_Success')

BaseCommunities_Success<-BaseCommunities_Success[1:200]

BaseCommunities_Success%>%
  map(CutModelDown)%>%
  map('web') %>%
  map(GetTL) %>% unlist-> NewTLs

BaseCommunities_Success %>%
  map(CutModelDown) -> CutDown

map_dbl(CutDown,'StartingNum')-> Sp
map(CutDown,'TL')%>% unlist-> TL
map(CutDown,'StartPoints')%>% unlist-> Eqm
map(CutDown,'K')%>% unlist-> K

## Body Size / TL
map(BaseCommunities_Success,function(Model){
return(Model$BMs[which(Model$Start>0)])
})%>% unlist-> BM

data.frame(BM, TL)%>%
  filter(TL>1,TL < 5)%>%
  ggplot(aes(col=factor(TL), fill= factor(TL), x=BM))+
  geom_density(alpha=0.5)+
  scale_x_log10()+
  geom_vline(xintercept =1, size=2 )+
 scale_color_discrete(name='Trophic Level')+
   scale_fill_discrete(name='Trophic Level')+
  labs(x='Body-mass', y='')

ggsave('Figures/Distribution of Body Masses.png', width = 6, height=4, dpi = 400)

```

### SI 2. Details of starting populations

```{r}
data.frame('NumberSp'=Sp)%>%
  ggplot(aes(x=Sp))+
  geom_bar()+
  scale_x_continuous(breaks=18:27, labels=18:27)+
  labs(x= 'Starting Community Size', y= 'Count') -> A

data.frame(NewTLs)%>%
  ggplot(aes(x=NewTLs))+
  geom_histogram(bins=20)+
  scale_x_continuous(breaks=c(1,2,2.5,3,3.5,4),
                     labels=c(1,2,2.5,3,3.5,4),
                     limits = c(0.9,4.1))+
  labs( x='Trophic Level Of Species', y='Number of Species') -> B
tail(sort(NewTLs),10)

data.frame(NewTLs, Eqm)%>%
  filter(Eqm > 0.01)%>%
  ggplot(aes(x=NewTLs, y=Eqm))+
  geom_point(alpha=0.1)+
  scale_y_log10(breaks=c(0.001, 0.01, 0.1, 1, 10, 100) ,
                labels=c(0.001, 0.01, 0.1, 1, 10, 100 ))+
  labs(y='Starting Equilibrium Biomass', x='Trophic Level Of Species') -> C


data.frame('C' =Eqm/K,  TL)%>%
  mutate(TLCat = ifelse(TL==1,'Producer', 'Consumer'  ))%>%
  ggplot(aes(x=C))+
  geom_histogram()+
  facet_wrap(~TLCat)+
  scale_x_continuous(breaks=c(0,0.25, 0.5, 0.75,1),
                     labels=c(0,0.25, 0.5, 0.75,1))+
  labs(x='Equilibrium Fraction of K', 
       y='Number of Species') -> D

plot_grid(A,B,C,D, labels = 'AUTO')
    
ggsave('Figures/CommunityInfoPlots.png', width = 8, height=8, dpi = 400)

```


## SI 5 Initial direction of response


```{r}
load(file='Output from Simulations/TIM_Models_Lots_Strong')

TIM_Models_Lots_Strong %>%
  map(TestInitialResponse) -> AllDirectionResponses

AllDirectionResponses%>% 
  unlist %>%
  table -> Dir_Table

Dir_Table / sum(Dir_Table)

Dir_Table[3]/(Dir_Table[1]+Dir_Table[3]) *100
```

## SI 7. Distribution of Population Shifts


```{r message=FALSE, warning=FALSE}
load('Output from Simulations/TIM_Models_Lots_Strong')
load('Output from Simulations/LS_TIM_Extinctions')

ExtractLosses<-function(model){
  starts<-  model$StartPoints    
  ends<- model$postExtEnd_List
  diffs <- matrix(NA, nrow=model$S, ncol=model$S)
  
  for( i in 1:model$S){
    try({
      diffs[i,] <- (ends[[i]] - starts)/starts
    }, silent = TRUE)
  }
  return(as.vector(diffs[!is.na(diffs)]))
}

LS_TIM_Extinctions%>%
  Unpacker%>%
  map(ExtractLosses)%>%
  unlist-> AllChanges

(AllChanges<0) %>% table

Extinctions <- AllChanges[AllChanges == -1]
Losses <-      AllChanges[AllChanges <0 & AllChanges!=-1]
Gains <-       AllChanges[AllChanges >0]

StdChanges<-data.frame(Value = c(-Extinctions, -Losses, Gains),
           Category = c(rep('Extinction',length(Extinctions)),
                        rep('Losses', length(Losses)), 
                        rep('Gains', length(Gains))))

table(StdChanges$Category)

StdChanges%>%
  filter(Category !='Extinction')%>%
  ggplot(aes(Value, fill=Category))+
  geom_histogram()+
  scale_x_log10()+
  facet_wrap(~Category, scales = 'free')+
  xlab('Proportional Change')+ ylab('')+
  scale_fill_brewer(palette = 'Set1')+
  guides(fill=FALSE)+

ggsave('Figures/SI7.png', dpi=400, height = 4, width = 9, units='in')

ExtractOverallLosses<-function(model){
  starts<-  model$StartPoints    
  ends<- model$postExtEnd_List
  diffs <- matrix(NA, nrow=model$S, ncol=1)

  for( i in 1:model$S){
    try({diffs[i,] <- (sum(ends[[i]]) - sum(starts))/sum(starts)}, silent = TRUE)}
  return(as.vector(diffs[!is.na(diffs)]))
}

LS_TIM_Extinctions%>%
  Unpacker%>%
  map(ExtractOverallLosses)%>%
  unlist-> OverallChange
table(OverallChange<0)

```



```{r eval=FALSE, include=FALSE}

### Motifs of Extinctions

## Additional analysis for R2

LS_TIM_unpacked<- Unpacker(LS_TIM_Extinctions)

MotifFreq<-function(run){
  
  Complete<- run$postExtEnd_List %>% map_lgl( function(x) {all(x!='DNF')})
  
  run$postExtEnd_List[Complete]%>%
    unlist%>%
    matrix(nrow = sum(Complete), byrow=TRUE)-> Results
  
  MotifFrequency <- data.frame('MotifFrequency' = rep(0,run$S ),
                               'n_Extinctions' = 0,
                               'OverallExtinctionFreq' = NA)
  
  for(ii in 1:sum(Complete)){
    
    target <- which(Complete)[ii]
    
    run$TIMlist%>%
      filter(i %in% which(rowSums(run$web)>1),
             k== target, 
             slope_ijk <0)%>%
      select(i)%>%
      unlist -> SuppressedResources
    
    MotifFrequency$OverallExtinctionFreq[target] <- mean(Results[ii,]==0)
    
    if(length(SuppressedResources>0)){
      MotifFrequency$MotifFrequency[target] <- length(SuppressedResources)
      MotifFrequency$n_Extinctions[target] <-  sum(Results[ii,SuppressedResources] ==0)
    }else{
      MotifFrequency$MotifFrequency[target] <-0
      MotifFrequency$n_Extinctions[target] <- 0
    }
  }
  return(MotifFrequency)
}

LS_TIM_unpacked %>%
  map_df(MotifFreq)-> OverallMotifFreqResults

OverallMotifFreqResults %>%
  mutate(MotifExtFreq = n_Extinctions /MotifFrequency )%>%
  filter(!is.na(OverallExtinctionFreq))%>%
  mutate(MotifExtFreq = ifelse(is.nan(MotifExtFreq),0,MotifExtFreq) )%>%
  colMeans()
```

# Session Info

```{r}

sessionInfo()
```